---
title: "lab_5_simone_albuquerque"
author: "Simone Albuquerque"
date: "October 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives
- Learn to parse dtes with 'Lubricate'
- Get counts of observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_tile()

```{r, include = FALSE}
# Read in data and attach packages

library(tidyverse)
library(janitor)
library(here)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>%  
  janitor::clean_names()
  
# view(lobster_abundance)
```

Use tidyr::uncount() function to conver our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

# view(lobster_tidy)
```

#### Exploratory Data Visualization 
We're only going to consider 'site' as our variable of interest

```{r}
ggplot(lobster_tidy, aes( x = site, y = size_mm)) +
         geom_jitter(aes(color = site), 
                     aplha = 0.5, 
                     width = 0.2)

# Histogram, one per site
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) + 
  facet_wrap(~site, scales = "free")
# made each graphs scale x,y independent to see trens within each site instead of comparing amongst sites

ggplot(lobster_tidy, aes(sample = size_mm)) + 
  geom_qq() +
  facet_wrap(~site)

#facet_wrap(~site) so that I could see if each site followed linear normal trend
```

#### They look normally distributed size_mm based on site.
 
- Central Limit Theorem tells us: since there are over 30 samples (hundreds), the means will be normally distributed (they would be normally distributed even if the stat_qq didnt show up normalized) and we can conduct a T test. 
 
 - This geom_qq tells us whether we trust the normal distribution of means as an honest representation of data
 
 #### Convert 'date' Column to class 'Date'
 
 We'll use the 'lubridate' package to convert to Date format, and then help us easily parse month & year 
 
```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

# view(lobster_date)
# class(lobster_date$date_new) to view if read as character or date
```
 
 Now let's parse year and month using lubridate::month() and lubridate::year():
 
```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), 
    obs_year = lubridate::year(date_new)
    
  )

view(lobster_parse_date)
class(lobster_parse_date$obs_month)

```
 
 Now let's fidn counts of observed lobsters based on different hierarchical listings
 First: let's count lobsters by year and month:
```{r}
lobster_ym <-  lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <-lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_y

lobster_site <-  lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
```
 
 If we want to create a summary table that contains statistics other than counts by group, it is easier to use group_by() + n()
 
```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE), 
    sd_size  = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )
lobster_summary
```
 #### Find Confidence Intervals
 
 Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples 
 
 - NULL hypothesis is that the mean is 0
```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
```
 
 #### Two - sample T- test to compare means
 
- Do I think I have enough evidence that these samples were drawn from populations that have different means?
 - the NULL hypothesis is that the samples are drawn from populations with the same mean
 
 We want to ask: Is there a significanct different in lobster lengths at Naples and Mohawk reefs? 
 
 We've done our necessary exploratory datato determine that a 2-sample t-test for means comparison is appropriate 
 
```{r}
napl_sample <-  lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <-  lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(mohk_sample, napl_sample)

mohk_napl_ttest
```
 
- It is so unlikely, based on p-value, that we would have drawn these two samples if they were drawn from pops with the same mean, so the more likely explanation is that they were drawn from populations with different means. .... 


##### There is a significant different in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p<0.001, alpha = 0.05)
 
-  Way more important are the qualitative different between the two means and differences in lobster lengths. The P-value is only interesting in that it leads up to more detailed understanding of data. This is simply a hypothesis test. 

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <-  t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic


# ~ means as a function of 
```

Here is the p-value: `r mn_ttest2$p.value`

Here is the t-statistic: `r mn_ttest2$statistic`

### Now: a heatmap! 

```{r}

lobster_ys <-  lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

 
 
 
